### 2024-01-26 Tables/ceoMD_overtime.tex

\begin{table}[h]
\footnotesize
\caption{Averages for CEO MD and non-MD Hospitals Over Time}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}ccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Beds} & \multicolumn{2}{c}{Penalized} & \multicolumn{2}{c}{HRRP Amount} & \multicolumn{2}{c}{Pneum. Mortality} & \multicolumn{2}{c}{Pneum. Readm.} & \multicolumn{2}{c}{HF Mort.} & \multicolumn{2}{c}{HF Readm.} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
\midrule
2008 & 72 & 16 &  &  &  &  & 10.8 & 12.5 &  &  & 10.9 & 10.5 &  & \\
2009 & 110 & 110 &  &  &  &  & 11.7 & 11.8 & 17.8 & 16.5 & 10.9 & 12.0 & 24.0 & 22.2\\
2010 & 109 & 32 &  &  &  &  & 11.9 & 11.3 & 18.3 & 17.8 & 11.5 & 10.3 & 24.7 & 25.0\\
2011 & 94 & 34 &  &  &  &  & 12.1 & 13.0 & 18.3 & 18.1 & 11.7 & 11.9 & 24.8 & 25.3\\
2012 & 97 & 247 & 0.29 & 0.50 & 51944 & 19601 & 12.2 & 12.2 & 18.3 & 18.3 & 12.0 & 12.8 & 24.5 & 24.6\\
\addlinespace
2013 & 68 & 116 & 0.32 & 0.50 & 65329 & 72226 & 12.1 & 11.9 & 17.4 & 18.1 & 12.1 & 12.2 & 22.9 & 23.5\\
2014 & 68 & 197 & 0.34 & 0.67 & 102671 & 134531 & 12.2 & 12.4 & 17.2 & 18.6 & 12.1 & 11.9 & 22.5 & 22.9\\
2015 & 82 & 100 & 0.40 & 0.50 & 123423 & 204336 & 11.7 & 10.6 & 16.7 & 17.5 & 12.0 & 10.6 & 21.8 & 22.6\\
\bottomrule
\end{tabular}
\end{table}