### 2024-01-26 Tables/cooMD_overtime.tex

\begin{table}[h]
\footnotesize
\caption{Averages for COO MD and non-MD Hospitals Over Time}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}ccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Beds} & \multicolumn{2}{c}{Penalized} & \multicolumn{2}{c}{HRRP Amount} & \multicolumn{2}{c}{Pneum. Mortality} & \multicolumn{2}{c}{Pneum. Readm.} & \multicolumn{2}{c}{HF Mort.} & \multicolumn{2}{c}{HF Readm.} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
\midrule
2008 & 71 & 25 &  &  &  &  & 10.8 & 11.0 &  &  & 10.9 & 10.0 &  & \\
2009 & 109 & 154 &  &  &  &  & 11.7 & 11.8 & 17.8 & 17.5 & 10.9 & 11.8 & 24.0 & 24.2\\
2010 & 110 & 48 &  &  &  &  & 11.8 & 12.5 & 18.3 & 17.5 & 11.5 & 11.5 & 24.7 & 24.8\\
2011 & 91 & 128 &  &  &  &  & 12.1 & 12.9 & 18.3 & 18.0 & 11.7 & 11.5 & 24.7 & 25.9\\
2012 & 99 & 112 & 0.30 & 0.33 & 51473 & 39378 & 12.2 & 14.2 & 18.2 & 19.0 & 12.0 & 11.9 & 24.5 & 24.5\\
\addlinespace
2013 & 69 & 87 & 0.32 & 0.33 & 65785 & 53204 & 12.0 & 13.9 & 17.4 & 18.7 & 12.2 & 11.1 & 22.9 & 24.6\\
2014 & 70 & 83 & 0.34 & 0.50 & 104078 & 79186 & 12.2 & 12.0 & 17.2 & 18.4 & 12.1 & 11.4 & 22.5 & 22.4\\
2015 & 82 & 137 & 0.39 & 0.75 & 128583 & 36959 & 11.6 & 10.6 & 16.8 & 17.0 & 12.0 & 12.1 & 21.9 & 21.9\\
\bottomrule
\end{tabular}
\end{table}