### 2024-01-26 Tables/presMD_overtime.tex

\begin{table}[h]
\footnotesize
\caption{Averages for President MD and non-MD Hospitals Over Time}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}ccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Beds} & \multicolumn{2}{c}{Penalized} & \multicolumn{2}{c}{HRRP Amount} & \multicolumn{2}{c}{Pneum. Mortality} & \multicolumn{2}{c}{Pneum. Readm.} & \multicolumn{2}{c}{HF Mort.} & \multicolumn{2}{c}{HF Readm.} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
\midrule
2008 & 72 & 25 &  &  &  &  & 10.8 & 11.0 &  &  & 10.9 & 10.7 &  & \\
2009 & 110 & 113 &  &  &  &  & 11.7 & 11.5 & 17.8 & 17.4 & 10.9 & 12.0 & 24.0 & 22.8\\
2010 & 110 & 100 &  &  &  &  & 11.9 & 11.4 & 18.3 & 17.9 & 11.5 & 11.6 & 24.7 & 24.8\\
2011 & 91 & 101 &  &  &  &  & 12.1 & 11.7 & 18.2 & 18.8 & 11.7 & 11.2 & 24.7 & 25.3\\
2012 & 98 & 111 & 0.27 & 0.46 & 57186 & 26860 & 12.2 & 12.2 & 18.2 & 18.4 & 12.0 & 11.9 & 24.5 & 24.9\\
\addlinespace
2013 & 68 & 86 & 0.30 & 0.50 & 63923 & 76254 & 12.0 & 12.3 & 17.3 & 17.8 & 12.2 & 11.8 & 22.8 & 23.3\\
2014 & 67 & 109 & 0.33 & 0.43 & 102995 & 110216 & 12.2 & 11.4 & 17.2 & 17.1 & 12.1 & 12.3 & 22.4 & 23.1\\
2015 & 82 & 89 & 0.39 & 0.45 & 127345 & 117249 & 11.7 & 11.5 & 16.8 & 16.6 & 12.0 & 12.4 & 21.8 & 22.0\\
\bottomrule
\end{tabular}
\end{table}