### 2024-01-26 Tables/Yearly_counts_name_data.tex

\begin{table}[h]
\footnotesize
\caption{Number of Hospitals with Relevant Leadership Positions}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}>{\centering\arraybackslash}p{.75cm}ccc}
\toprule
Year & Total & Any MD & CEO & MD CEO & Pres & MD Pres & CEO or Pres & MD CEO or Pres & CFO & MD CFO & CMO & MD CMO & COO & MD COO\\
\midrule
2008 & 109 & 8 & 70 & 2 & 55 & 4 & 87 & 6 & 64 & 1 & 0 & 0 & 14 & 1\\
2009 & 167 & 25 & 105 & 5 & 96 & 11 & 137 & 13 & 89 & 1 & 13 & 8 & 46 & 6\\
2010 & 251 & 46 & 111 & 5 & 103 & 22 & 159 & 24 & 102 & 3 & 22 & 10 & 44 & 4\\
2011 & 225 & 40 & 130 & 8 & 115 & 17 & 178 & 24 & 115 & 1 & 19 & 12 & 48 & 6\\
2012 & 295 & 61 & 234 & 6 & 201 & 37 & 281 & 39 & 216 & 1 & 43 & 25 & 71 & 6\\
\addlinespace
2013 & 233 & 42 & 184 & 8 & 169 & 21 & 226 & 26 & 169 & 0 & 29 & 17 & 50 & 3\\
2014 & 181 & 34 & 137 & 3 & 123 & 12 & 171 & 14 & 132 & 0 & 23 & 16 & 36 & 2\\
2015 & 291 & 64 & 219 & 9 & 202 & 27 & 277 & 33 & 210 & 0 & 44 & 28 & 61 & 4\\
\bottomrule
\end{tabular}
\end{table}