### 2024-01-26 Tables/anyMD_overtime.tex

\begin{table}[h]
\footnotesize
\caption{Averages for MD and non-MD Hospitals Over Time}
\centering
\begin{tabular}[t]{>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}>{\centering\arraybackslash}p{.9cm}ccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Beds} & \multicolumn{2}{c}{Penalized} & \multicolumn{2}{c}{HRRP Amount} & \multicolumn{2}{c}{Pneum. Mortality} & \multicolumn{2}{c}{Pneum. Readm.} & \multicolumn{2}{c}{HF Mort.} & \multicolumn{2}{c}{HF Readm.} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15}
Year & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD & No MD & MD\\
\midrule
2008 & 74 & 23 &  &  &  &  & 10.8 & 11.5 &  &  & 10.9 & 10.5 &  & \\
2009 & 108 & 124 &  &  &  &  & 11.7 & 11.4 & 17.9 & 17.2 & 10.9 & 11.4 & 24.1 & 23.4\\
2010 & 107 & 115 &  &  &  &  & 11.9 & 11.6 & 18.4 & 18.0 & 11.5 & 11.6 & 24.7 & 24.9\\
2011 & 90 & 101 &  &  &  &  & 12.1 & 12.1 & 18.2 & 18.4 & 11.7 & 11.5 & 24.7 & 25.3\\
2012 & 89 & 140 & 0.26 & 0.47 & 60412 & 31043 & 12.1 & 12.5 & 18.2 & 18.4 & 12.0 & 12.1 & 24.5 & 24.6\\
\addlinespace
2013 & 67 & 79 & 0.29 & 0.46 & 62853 & 74043 & 12.0 & 12.4 & 17.3 & 17.9 & 12.2 & 12.1 & 22.8 & 23.4\\
2014 & 70 & 74 & 0.34 & 0.35 & 104178 & 101575 & 12.1 & 12.4 & 17.2 & 17.5 & 12.1 & 12.1 & 22.4 & 22.8\\
2015 & 85 & 73 & 0.41 & 0.37 & 134604 & 92287 & 11.7 & 11.5 & 16.7 & 16.8 & 12.0 & 12.0 & 21.8 & 21.9\\
\bottomrule
\end{tabular}
\end{table}