### 2024-04-23 Tables/MD_noMD_uncompCMI_synth.tex

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_uncompCMI_synth}Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Uncomp. Care (Millions)} & \multicolumn{1}{c}{Case Mix Index} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
 & (1) & (2)\\
\midrule
NP w/ MD Executive x Post Penalty & 3.12 & 0.03$^{**}$\\
 & (3.04) & (0.01)\\
\textbf{Comparison Group:} &  & \\
NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
Observations & 625 & 1635\\
\bottomrule
\end{tabular}
\end{table}