### 2024-04-23 Tables/MD_noMD_uncompCMI_matchsample.tex

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_uncomp_CMI_matchsample} Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Regression Results, Matched Sample}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Uncomp. Care (Millions) & Case Mix Index\\  
                                        & (1)                     & (2)\\  
      \midrule 
      NP w/ MD Executive x Post Penalty & 10.53                   & 0.03$^{**}$\\   
                                        & (7.39)                  & (0.01)\\   
      \textbf{Comparison Group:}        &                         & \\  
      NP w/out MD Executive             & $\checkmark$            & $\checkmark$\\   
       \\
      Observations                      & 893                     & 996\\  
      \bottomrule
   \end{tabular}
\end{table}