### 2024-04-23 Tables/forprofit_synth.tex

\begin{table}[ht!]

\caption{\label{tab:forprofit_synth} Comparison to For-Profit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For Profit x Post Penalty & -0.07 & -0.32$^{**}$ & -0.03 & -0.01 & -0.17 & 0.02\\
 & (0.04) & (0.12) & (0.06) & (0.04) & (0.11) & (0.05)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
\\
Observations & 15,595 & 2,900 & 4,705 & 15,515 & 2,895 & 4,690\\
\bottomrule
\end{tabular}
\end{table}