### 2024-04-25 Tables/MD_noMD_readmort_synth_2012MD.tex

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_synth}Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
 & (1) & (2)\\
\midrule
NP w/ MD Executive x Post Penalty & -0.09 & -0.1\\
 & (0.07) & (0.06)\\
\textbf{Comparison Group:} &  & \\
NP w/out MD Executive & $\checkmark$ & $\checkmark$\\
Observations & 3925 & 3910\\
\bottomrule
\end{tabular}
\end{table}