### 2024-04-25 Tables/MD_noMD_uncompCMI_fullsample_2012MD.tex

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_uncomp_CMI_fullsample} Comparison of MD vs. non-MD Nonprofit Uncompensated Care and CMI Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Uncomp. Care (Millions) & Case Mix Index\\  
                                        & (1)                     & (2)\\  
      \midrule 
      NP w/ MD Executive x Post Penalty & -2.36                   & 0.01$^{**}$\\   
                                        & (2.75)                  & (0.01)\\   
      \textbf{Comparison Group:}        &                         & \\  
      NP w/out MD Executive             & $\checkmark$            & $\checkmark$\\   
       \\
      Observations                      & 2,483                   & 3,046\\  
      \bottomrule
   \end{tabular}
\end{table}