### 2024-04-25 Tables/forprofit_uncompCMI_synth_everMD.tex

\begin{table}[ht!]

\caption{\label{tab:forprofit_uncompCMI_synth} Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncomp. Care (Millions)} & \multicolumn{3}{c}{Case Mix Index} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For Profit x Post Penalty & 10.52$^{***}$ & 9.69$^{**}$ & 9.61$^{***}$ & 0 & 0 & 0\\
 & (2.53) & (2.96) & (2.28) & (0.01) & (0.01) & (0.01)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
\addlinespace
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
Observations & 4940 & 1895 & 4305 & 12320 & 4250 & 10715\\
\bottomrule
\end{tabular}
\end{table}