### 2024-04-30 Tables/MD_noMD_readmort_fullsample_2012MD.tex
update tables

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_fullsample} Comparison of MD vs. non-MD Nonprofit Readmission and Mortality Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Weighted Avg. Readmission Rate & Weighted Avg. Mortality Rate\\  
                                        & (1)                            & (2)\\  
      \midrule 
      NP w/ MD Executive x Post Penalty & -0.27$^{***}$                  & -0.11\\   
                                        & (0.10)                         & (0.10)\\   
      \textbf{Comparison Group:}        &                                & \\  
      NP w/out MD Executive             & $\checkmark$                   & $\checkmark$\\   
       \\
      Observations                      & 1,883                          & 1,880\\  
      \bottomrule
   \end{tabular}
\end{table}