### 2024-04-30 Tables/MD_noMD_readmort_decomp_fullsample.tex
add decomposition tables

\begin{table}[htbp]
   \caption{\label{tab:MD_noMD_readmort_decomp_fullsample} Decomposition of Readmission and Mortality Results for Nonprofits}
   \bigskip
   \centering
   \begin{tabular}{lcccc}
      \toprule
       & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate}\\
                                       & (1)           & (2)           & (3)           & (4)\\  
      \midrule 
      Ever MD Executive x Post Penalty & -0.03         &               & -0.02         &   \\   
                                       & (0.07)        &               & (0.07)        &   \\   
      MD Executive in 2012             &               & -0.27$^{***}$ &               & -0.11\\   
                                       &               & (0.10)        &               & (0.10)\\   
      \textbf{Comparison Group:}       &               &               &               & \\  
      Never MD                         & $\checkmark$  &               & $\checkmark$  &  \\  
      Has MD not in 2012               &               & $\checkmark$  &               & $\checkmark$\\   
       \\
      Observations                     & 4,095         & 1,883         & 4,083         & 1,880\\  
      \bottomrule
   \end{tabular}
\end{table}