### 2024-04-30 Tables/MD_noMD_uncompCMI_decomp_synth.tex

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_uncompCMI_decomp_synth}Decomposition of Uncompensated Care and CMI for Nonprofits Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Uncompensated Care} & \multicolumn{2}{c}{Case Mix Index} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
 & (1) & (2) & (1) & (2)\\
\midrule
Ever MD Executive x Post 2012 & 0.01 &  & 0.01$^{*}$ & \\
 & (0.01) &  & (0) & \\
2012 MD Executive x Post 2012 &  & -0.03 &  & 0.01\\
 &  & (0.03) &  & (0.01)\\
 &  &  &  \vphantom{1} & \\
\addlinespace
\textbf{Comparison Group:} &  &  &  & \\
Never MD & $\checkmark$ &  & $\checkmark$ & \\
Has MD not in 2012 &  & $\checkmark$ &  & $\checkmark$\\
 &  &  &  & \\
Observations & 1195 & 635 & 3015 & 1595\\
\bottomrule
\end{tabular}
\end{table}