### 2024-05-01 Tables/MD_noMD_readmort_fullsample.tex
make tables/figures look better and more consistent

\begin{table}[ht!]
   \caption{\label{tab:MD_noMD_readmort_fullsample} Non-MD vs. MD Executive Readmission and Mortality Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                       & Weighted Avg. Readmission Rate & Weighted Avg. Mortality Rate\\  
                                       & (1)                            & (2)\\  
      \midrule 
      NP w/out MD Exec x Post Programs & -0.40$^{***}$                  & -0.19\\   
                                       & (0.14)                         & (0.15)\\   
      \textbf{Comparison Group:}       &                                & \\  
      NP w/ MD Exec                    & $\checkmark$                   & $\checkmark$\\   
       \\
      Observations                     & 2,471                          & 2,462\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
\end{table}