### 2024-05-01 Tables/MD_noMD_uncompCMI_fullsample.tex

\begin{table}[ht!]
   \caption{\label{tab:MD_noMD_uncompCMI_fullsample} Non-MD vs. MD Executive Uncompensated Care and CMI Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                       & Uncompensated Care & Case Mix Index\\  
                                       & (1)                & (2)\\  
      \midrule 
      NP w/out MD Exec x Post Programs & 0.01               & 0.03$^{***}$\\   
                                       & (0.01)             & (0.01)\\   
      \textbf{Comparison Group:}       &                    & \\  
      NP w/ MD Executive               & $\checkmark$       & $\checkmark$\\   
       \\
      Observations                     & 1,440              & 1,654\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:forprofit} with readmission and mortality rates as outcome variables.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1\\
   Uncompensated care measured as percent of total operating expenses.
\end{table}