### 2024-05-01 Tables/MD_noMD_readmort_synth.tex
Updates from Overleaf

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_synth}Comparison of MD vs. non-MD Nonprofit Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3}
 & (1) & (2)\\
\midrule
NP w/out MD Exec x Post Penalty & -0.28$^{*}$ & -0.19\\
 & (0.12) & (0.11)\\
\textbf{Comparison Group:} &  & \\
NP w/ MD Executive & $\checkmark$ & $\checkmark$\\
Observations & 2335 & 2325\\
\bottomrule
\end{tabular}
\end{table}