### 2024-05-01 Tables/MD_noMD_readmort_decomp_fullsample.tex
Update tables AGAIN

\begin{table}[ht!]
   \caption{\label{tab:MD_noMD_readmort_decomp_fullsample} Decomposition of Readmission and Mortality Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcccc}
      \toprule
       & \multicolumn{2}{c}{Weighted Avg. Readmission Rate} & \multicolumn{2}{c}{Weighted Avg. Mortality Rate}\\
                                   & (1)           & (2)           & (3)           & (4)\\  
      \midrule 
      Ever MD Exec x Post Programs & -0.03         &               & -0.02         &   \\   
                                   & (0.07)        &               & (0.07)        &   \\   
      MD Exec in 2012              &               & -0.27$^{***}$ &               & -0.11\\   
                                   &               & (0.10)        &               & (0.10)\\   
      \textbf{Comparison Group:}   &               &               &               & \\  
      Never MD                     & $\checkmark$  &               & $\checkmark$  &  \\  
      Has MD not in 2012           &               & $\checkmark$  &               & $\checkmark$\\   
       \\
      Observations                 & 4,095         & 1,883         & 4,083         & 1,880\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:decomp} with readmission and mortality rates as outcome variables.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
\end{table}