### 2024-05-01 Tables/MD_noMD_readmort_synth.tex

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_readmort_synth}Non-MD vs. MD Executive Readmission and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Weighted Avg. Readmission Rate} & \multicolumn{1}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2)\\
\midrule
No MD Exec x Post Programs & -0.28$^{*}$ & -0.19\\
 & (0.13) & (0.11)\\
\textbf{Comparison Group:} &  & \\
NP w/ MD Exec & $\checkmark$ & $\checkmark$\\
 &  & \\
\addlinespace
Observations & 2,335 & 2,325\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.}\\
\multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{3}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}