### 2024-05-01 Tables/forprofit_uncompCMI_synth.tex

\begin{table}[ht!]

\caption{\label{tab:forprofit_uncompCMI_synth}Comparison to For-Profit Uncompensated Care and CMI Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Uncompensated Care} & \multicolumn{3}{c}{Case Mix Index} \\
 & (1) & (2) & (3) & (1) & (2) & (3)\\
\midrule
For-Profit x Post Programs & 0.06$^{***}$ & 0.08$^{***}$ & 0.06$^{***}$ & 0 & 0.02$^{**}$ & 0\\
 & (0.01) & (0.01) & (0.01) & (0.01) & (0.01) & (0.01)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
All Nonprofits & $\checkmark$ &  &  & $\checkmark$ &  & \\
Nonprofits w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
Nonprofits w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
 &  &  &  &  &  & \\
Observations & 4,935 & 1,330 & 1,815 & 12,320 & 2,880 & 4,065\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with uncompensated care and CMI as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\multicolumn{7}{l}{\textsuperscript{} Uncompensated care measured as percent of total operating expenses.}\\
\end{tabular}
\end{table}