### 2024-05-06 Tables/forprofit_readmort_synth.tex
Add individual synthetic diff in diff plots

\begin{table}[ht!]

\caption{\label{tab:forprofit_readmort_synth}Comparison to For-Profit Readmission and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Weighted Avg. Readmission Rate} & \multicolumn{3}{c}{Weighted Avg. Mortality Rate} \\
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\midrule
For-Profit x Post Programs & -0.03 & -0.32$^{**}$ & -0.03 & -0.01 & -0.17 & 0.02\\
 & (0.05) & (0.12) & (0.06) & (0.04) & (0.11) & (0.05)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
In-Sample NFPs & $\checkmark$ &  &  & $\checkmark$ &  & \\
NFPs w/ MD &  & $\checkmark$ &  &  & $\checkmark$ & \\
\addlinespace
NFPs w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$\\
 &  &  &  &  &  & \\
Observations & 6,560 & 2,900 & 4,705 & 6,540 & 2,895 & 4,690\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with readmission and mortality rates as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}