### 2024-05-14 Tables/MD_noMD_readmort_fullsample.tex
change tables and graphs to be adjusted for case mix index

\begin{table}[ht!]
   \caption{\label{tab:MD_noMD_readmort_fullsample} Non-MD vs. MD Executive Readmission and Mortality Regression Results}
   \bigskip
   \centering
   \begin{tabular}{lcc}
      \toprule
                                        & Weighted Avg. Readmission Rate & Weighted Avg. Mortality Rate\\  
                                        & (1)                            & (2)\\  
      \midrule 
      NFP w/out MD Exec x Post Programs & -0.35$^{**}$                   & -0.14\\   
                                        & (0.16)                         & (0.15)\\   
      \textbf{Comparison Group:}        &                                & \\  
      NFP w/ MD Exec                    & $\checkmark$                   & $\checkmark$\\   
       \\
      Observations                      & 1,648                          & 1,648\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:clinical} with readmission and mortality rates as outcome variables.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
\end{table}