### 2024-05-30 Tables/MD_noMD_readmort_condition_synth.tex
add tables for specific conditions and dsecriptive stats about leadership team changes

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_readmort_condition_synth}Non-MD vs. MD Executive Readmission and Mortality Synthetic DiD Results, All Conditions}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\underline{ }} & \multicolumn{3}{c}{\underline{Readmission Rates}} & \multicolumn{3}{c}{\underline{Mortality Rates}} \\
 & AMI & HF & Pneum. & AMI & HF & Pneum.\\
\midrule
NFP w/out MD Exec x Post Program & -0.19 & -0.18 & -0.4$^{*}$ & 0.09 & -0.1 & -0.3\\
 & (0.19) & (0.2) & (0.17) & (0.2) & (0.15) & (0.19)\\
\textbf{Comparison Group:} &  &  &  &  &  & \\
NFP w/ MD Exec & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
 &  &  &  &  &  & \\
\addlinespace
Observations & 1,135 & 1,625 & 1,625 & 1,295 & 1,625 & 1,625\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Results from Equation with condition specific readmission and mortality rates as outcome variables.}\\
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}