### 2024-05-30 Tables/md_readmort_condition_fullsample.tex

\begin{table}[ht!]
   \caption{\label{tab:forprofit_mort_conditions_fullsample} Comparison to For-Profit Mortality Regression Results, All Conditions}
   \bigskip
   \centering
   \begin{tabular}{lcccccc}
      \toprule
                                 & Heart Attack & Heart Failure & Pneumonia     & Heart Attack & Heart Failure & Pneumonia\\  
       & \multicolumn{3}{c}{Readmission Rates} & \multicolumn{3}{c}{Mortality Rates} \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7}
                                 & (1)          & (2)           & (3)           & (4)          & (5)           & (6)\\  
      \midrule 
      For-Profit x Post Programs & -0.20        & -0.20         & -0.48$^{***}$ & -0.08        & -0.03         & -0.33\\   
                                 & (0.20)       & (0.20)        & (0.15)        & (0.21)       & (0.18)        & (0.23)\\   
      \textbf{Comparison Group:} &  &  &  &  &  &  &  &  & \\
      In-Sample NFPs & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$ &  & \\
      NFP w/ MD &  & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$ & \\
      NFPs w/out MD &  &  & $\checkmark$ &  &  & $\checkmark$ &  &  & $\checkmark$\\
       \\
      Observations               & 1,293        & 2,342         & 2,467         & 1,514        & 2,288         & 2,459\\  
      \bottomrule
   \end{tabular}

   \par \raggedright 
   Results from Equation \ref{eq:forprofit} for mortality rates by condition.\\
   Standard errors are clustered at the hospital level.\\
   Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1
\end{table}