### 2024-05-30 Tables/MD_noMD_readmort_decomp_synth.tex

\begin{table}[ht!]

\caption{\label{tab:MD_noMD_readmort_decomp_synth}Decomposition of Readmissions and Mortality Synthetic DiD Results}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
\addlinespace
 & (1) & (2) & (3) & (4)\\
\midrule
Never MD Exec x Post Programs & 0 &  & -0.06 & \\
 & (0.06) &  & (0.06) & \\
Ever MD (not in 2012) &  & -0.19$^{*}$ &  & -0.1\\
 \hspace{3mm} x Post Programs &  & (0.09) &  & (0.08)\\
 &  &  &  & \\
\addlinespace
\textbf{Comparison Group:} &  &  &  & \\
Ever MD Exec & $\checkmark$ &  & $\checkmark$ & \\
Has MD Exec 2012 &  & $\checkmark$ &  & $\checkmark$\\
 &  &  &  & \\
Observations & 3,925 & 1,845 & 3,910 & 1,840\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Results from Equation \ref{eq:decomp} with readmission and mortality rates as outcome variables.}\\
\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}