### 2024-08-21 Tables/main_twfe_match_tab.tex

\begin{table}[ht!]

\caption{\label{tab:main_twfe_match}Effect of Clinically Trained Executive, TWFE and Matched Sample}
\centering
\begin{tabular}[t]{lcccl}
\toprule
\multicolumn{1}{c}{\underline{ }} & \multicolumn{2}{c}{\underline{TWFE}} & \multicolumn{2}{c}{\underline{Matched Sample}} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} & \multicolumn{1}{c}{Read} & \multicolumn{1}{c}{Mort} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5}
 & (1) & (2) & (3) & (4)\\
\midrule
Ever Clinical Exec x Post Program & 0.42$^{**}$ & 0.11 & 0.65$^{*}$ & 0.03\\
 & (0.15) & (0.15) & (0.3) & (0.4)\\
 &  &  &  & \\
Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Observations & 2,387 & 2,381 & 1,025 & 1,023\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{5}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}