### 2024-08-21 Tables/condition_synthdid_table.tex
Updates from Overleaf

\begin{table}[ht!]

\caption{\label{tab:results_by_condition}Effect of Clinically Trained Executive on Condition-Specific Readmission and Mortality Rates}
\centering
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Readmission Rates} & \multicolumn{3}{c}{Mortality Rates} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
\addlinespace[0.3em]
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} & \multicolumn{1}{c}{AMI} & \multicolumn{1}{c}{HF} & \multicolumn{1}{c}{Pneum.} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\midrule
Always Clinical Exec x Post Program & 0.25 & 0.14 & 0.46$^{**}$ & -0.13 & 0.03 & 0.3\\
 & (0.18) & (0.18) & (0.14) & (0.19) & (0.13) & (0.19)\\
 &  &  &  &  &  & \\
Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Observations & 1,135 & 1,640 & 1,635 & 1,300 & 1,640 & 1,635\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}