### 2024-08-23 Tables/fp_always_estimators_mort_tab.tex
add tables for appendix, various estimators

\begin{table}[ht!]

\caption{\label{tab:forprofit_mort_estimators}Compare For-Profit to Types of Leadership (Mortality Rates), Various Estimations}
\centering
\begin{tabular}[t]{lcccccl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Compare to Always Clinical Exec} & \multicolumn{3}{c}{Compare to Never Clinical Exec} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} & \multicolumn{1}{c}{TWFE} & \multicolumn{1}{c}{Matched} & \multicolumn{1}{c}{SC} \\
\cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7}
 & (1) & (2) & (3) & (4) & (5) & (6)\\
\midrule
For-Profit x Post Program & 0.03 & 0.09 & -0.16 & 0.06 & 0.05 & 0.03\\
 & (0.14) & (0.45) & (0.2) & (0.07) & (0.11) & (0.1)\\
 &  &  &  &  &  & \\
Hospital FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Year FE & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Observations & 2,660 & 660 & 2,660 & 3,820 & 1,790 & 3,820\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{7}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}