### 2024-11-06 Slides/JobTalk.tex
Updates from Overleaf

\documentclass[notes,11pt, aspectratio=169]{beamer}

\usepackage{pgfpages}
% These slides also contain speaker notes. You can print just the slides,
% just the notes, or both, depending on the setting below. Comment out the want
% you want.
\setbeameroption{hide notes} % Only slide
%\setbeameroption{show only notes} % Only notes
%\setbeameroption{show notes on second screen=right} % Both

\usepackage{helvet}
\usepackage[default]{lato}
\usepackage{array}

\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{Paper/papercitations.bib}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}
\usepackage{import}

\usepackage{changepage}
\usepackage{appendixnumberbeamer}
\newcommand{\beginbackup}{
   \newcounter{framenumbervorappendix}
   \setcounter{framenumbervorappendix}{\value{framenumber}}
   \setbeamertemplate{footline}
   {
     \leavevmode%
     \hline
     box{%
       \begin{beamercolorbox}[wd=\paperwidth,ht=2.25ex,dp=1ex,right]{footlinecolor}%
%         \insertframenumber  \hspace*{2ex} 
       \end{beamercolorbox}}%
     \vskip0pt%
   }
 }
\newcommand{\backupend}{
   \addtocounter{framenumbervorappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumbervorappendix}} 
}

\setbeamertemplate{blocks}[rounded]
\setbeamercolor{block title}{bg=eggplant!70, fg=white}
\setbeamercolor{block body}{bg=eggplant!20}

\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{booktabs}

% These are my colors -- there are many like them, but these ones are mine.
\definecolor{sage}{RGB}{102,153,102}
\definecolor{cambridgeblue}{RGB}{104, 166, 145}
\definecolor{yellow}{RGB}{255,173,1}
\definecolor{purple}{RGB}{153,102,153}
\definecolor{ashgray}{RGB}{191, 211, 193}
\definecolor{eggplant}{RGB}{105, 79, 93}
\definecolor{silver}{RGB}{192, 197, 193}
\definecolor{darkcyan}{RGB}{78, 135, 140}
\definecolor{lightcyan}{RGB}{201, 228, 231}
\definecolor{columbiablue}{RGB}{187, 213, 237}
\definecolor{lightblue}{RGB}{184, 219, 217}
\definecolor{earthyellow}{RGB}{234, 180, 100}

\setbeamercolor{mycolor}{fg=white,bg=sage}

% colors for diagrams
\definecolor{diagramtan}{RGB}{225,190,106}
\definecolor{diagramteal}{RGB}{64, 176, 166}
\definecolor{diagrampurple}{RGB}{170, 131, 239}
\definecolor{diagramred}{RGB}{146, 79, 79}

\hypersetup{
  colorlinks=false,
  linkbordercolor = {white},
  linkcolor = {sage}
}

\newcommand{\btVFill}{\vskip0pt plus 1filll}

%% I use a beige off white for my background
\definecolor{MyBackground}{RGB}{255,253,218}

%% Uncomment this if you want to change the background color to something else
%\setbeamercolor{background canvas}{bg=MyBackground}

%% Change the bg color to adjust your transition slide background color!
\newenvironment{transitionframe}{
  \setbeamercolor{background canvas}{bg=lightcyan}
  \begin{frame}[plain,noframenumbering]}{
    \end{frame}
}

\setbeamercolor{frametitle}{fg=black, bg=lightcyan}
\setbeamercolor{title}{fg=black}
\setbeamertemplate{footline}{%
  \raisebox{5pt}{\makebox[\paperwidth]{\hfill\makebox[15pt]{\scriptsize\insertframenumber}}}}
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\bullet$}
\setbeamercolor{itemize item}{fg=lightcyan}
\setbeamercolor{itemize subitem}{fg=lightcyan}
\setbeamercolor{enumerate item}{fg=black}
\setbeamercolor{enumerate subitem}{fg=black}
\setbeamercolor{button}{bg=eggplant!20,fg=black}

\setbeamertemplate{frametitle}
{
    \nointerlineskip
    \begin{beamercolorbox}[sep=0.3cm,ht=2em,wd=\paperwidth]{frametitle}
        \vbox{}\vskip-2ex%
        \strut\insertframetitle\strut
        \vskip-0.8ex%
    \end{beamercolorbox}
}

% If you like road maps, rather than having clutter at the top, have a roadmap show up at the end of each section 
% (and after your introduction)
% Uncomment this is if you want the roadmap!
% \AtBeginSection[]
% {
%    \begin{frame}
%        \frametitle{Roadmap of Talk}
%        \tableofcontents[currentsection]
%    \end{frame}
% }

\setbeamercolor{section in toc}{fg=sage}
\setbeamercolor{subsection in toc}{fg=sage}
\setbeamersize{text margin left=1em,text margin right=1em}

\newenvironment{wideitemize}{\itemize\addtolength{\itemsep}{10pt}}{\enditemize}

\usepackage{environ}
\NewEnviron{videoframe}[1]{
  \begin{frame}
    \vspace{-8pt}
    \begin{columns}[onlytextwidth, T] % align columns
      \begin{column}{.58\textwidth}
        \begin{minipage}[t][\textheight][t]
          {\dimexpr\textwidth}
          \vspace{8pt}
          \hspace{4pt} {\Large \sc \textcolor{blue}{#1}}
          \vspace{8pt}

          \BODY
        \end{minipage}
      \end{column}%
      \hfill%
      \begin{column}{.42\textwidth}
        \colorbox{green!20}{\begin{minipage}[t][1.2\textheight][t]
            {\dimexpr\textwidth}
            Face goes here
          \end{minipage}}
      \end{column}%
    \end{columns}
  \end{frame}
}

\title[]{\textcolor{sage}{Does Hospital Leadership Matter? \newline
Evidence from Pay-for-Performance}}

\author[]{Hanna Glenn}
\date{}

\begin{document}

%%% TIKZ STUFF
\tikzset{   
        every picture/.style={remember picture,baseline},
        every node/.style={anchor=base,align=center,outer sep=1.5pt},
        every path/.style={thick},
        }
\newcommand\marktopleft[1]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-a) at (-.3em,.3em) {};%
}
\newcommand\markbottomright[2]{%
    \tikz[overlay,remember picture] 
        \node (marker-#1-b) at (0em,0em) {};%
}
\tikzstyle{every picture}+=[remember picture] 
\tikzstyle{mybox} =[draw=black, very thick, rectangle, inner sep=10pt, inner ysep=20pt]
\tikzstyle{fancytitle} =[draw=black,fill=red, text=white]
%%%% END TIKZ STUFF

% Title Slide
\begin{frame}[plain]
% Background image
\begin{tikzpicture}[remember picture,overlay]
    \node
[
    above=0.5cm,
    align=center,
    fill=lightcyan,
    rounded corners,
    inner xsep=15pt,
    inner ysep=10pt, 
    minimum width=0.9\textwidth,
    text width=0.6\textwidth
] (title) at (current page.center)
{
    \LARGE Does Hospital Leadership Matter?  \\[5pt]
    \normalsize Evidence from Pay-for-Performance
};
% Author 
\node[below=1cm] (author) at (title.south){Hanna Glenn, Emory University};
% Date
\node[below=0.25cm] (date) at (author.south){\today};
\end{tikzpicture}

\end{frame}

\section{Intro}

\begin{frame}{Firm Leadership}
\centering
    \begin{tikzpicture}
        % First image (top-left)
        \node[anchor=north west] (img1) at (0, 3) {\includegraphics[width=0.55\textwidth]{Graphics/theroleleadershiphasincompanyculture.png}};
        \draw[line width=.45mm, lightcyan] (img1.north west) rectangle (img1.south east);

        % Second image (slightly down and right of the first image)
        \node[anchor=north west] (img2) at (0.28, 0.58) {\includegraphics[width=0.5\textwidth]{Graphics/linkingbusinessstrategyandleadershipforbetteroutcomes.png}};
        \draw[line width=.45mm, lightcyan] (img2.north west) rectangle (img2.south east);

        % Fourth image (bottom-right)
        \node[anchor=north west] (img4) at (0, -2) {\includegraphics[width=0.55\textwidth]{Graphics/theroleofleadershipinorganizationalsuccess.png}};
        \draw[line width=.45mm, lightcyan] (img4.north west) rectangle (img4.south east);

    \end{tikzpicture}
\end{frame}

\begin{frame}{Leadership Characteristics}

\vspace{1mm}

\textbf{Gender}
\begin{itemize}
    \item \small More female board members $\implies$ lower firm value \scriptsize(\cite{ahern2012changing}) \small
    \item \small More female board members $\implies$ less employment reduction \scriptsize (\cite{matsa2013female}) \small
    \item female executives affect the female wage distribution \scriptsize (\cite{flabbi2019female}) \small
\end{itemize}

\vspace{2mm}

\textbf{Background}
\begin{itemize}
    \item \small Firms pay more for general background CEOs \scriptsize (\cite{custodio2013generalists}) \small
    \item \small CEOs w/ prior experience do better in M\&A \scriptsize(\cite{custodio2013ceos}) \small
    \item \small Military experience CEOs are less aggressive in R\&D and investing \scriptsize (\cite{benmelech2015military})
\end{itemize}

\vspace{2mm}

\textbf{Age}
\begin{itemize}
    \item \small Young male CEOs more aggressive in M\&A \scriptsize (\cite{levi2010deal})
\end{itemize}

\btVFill\pause

\begin{block}{\normalsize Gaps in Literature}
    \small 
    \begin{enumerate}
        \item Nonprofit firms: reason to believe these findings might not inform leadership in other settings
        \item Direct responses to policy changes
    \end{enumerate}
\end{block}
\end{frame}

\begin{frame}{Hospitals \& Health Care Policy Goals}

\begin{block}{This Paper}
    Examine the role of leadership in nonprofit hospitals responding to policy incentives 
\end{block}

\vspace{5mm}

\begin{enumerate}
    \item \textbf{Hospital care makes up a third of health care spending} \scriptsize (\cite{cms2024nhe}) \normalsize

\vspace{2mm}

    \begin{itemize}
        \item Nonprofits drive the hospital industry \scriptsize (\cite{ASPE_2023}) \normalsize
    \end{itemize}

\vspace{5mm}

    \item \textbf{Policies often aim to improve the value of health care delivery}

\vspace{2mm}

    \begin{wideitemize}
        \item It is unclear how hospital leaders view value
        \item Or whether they might play a role in achieving better value
    \end{wideitemize}

    \end{enumerate}
\end{frame}

\begin{frame}{Research Question}
    \centering
    \large
    Does having executives with clinical training affect response to pay-for-performance incentives of nonprofit hospitals in the US?
\end{frame}

\begin{frame}{Why Clinical Training?}

\textbf{Do physicians make good firm leaders?}

\vspace{2mm}

\begin{wideitemize}
    \item Physicians can bring a unique combination of clinical and administration expertise that can potentially lead to better quality \scriptsize (\cite{Stajduhar_2023}, \cite{Ahmed_2022}) \normalsize

    \vspace{2mm}

    \item Not always trained in business management, which can lead to poor practices and lower quality \scriptsize (\cite{HarvardBusinessReview2018})
\end{wideitemize}

\vspace{10mm}

\textbf{Clinical training is potentially correlated with how the leader values quality}
\end{frame}

\begin{frame}{Preview of Findings}

\textbf{Non-clinical executive teams respond more drastically to financial incentives on quality compared to clinical executive teams}

\vspace{2mm}

\begin{wideitemize}
    \item Non-clinical reduce readmission by 2\% more than clinical
    \item Non-clinical executive teams behave like profit-driven hospitals
\end{wideitemize}
\end{frame}

\begin{frame}{Contribution}
    \textbf{Hospital Executives/Board of Directors}
    \begin{itemize}
        \item Physician board members $\implies$ lower public donations in nonprofit hospitals \scriptsize (\cite{brickley2010board}) \normalsize
        \item No correlation between CEOs and hospital production in England \scriptsize (\cite{janke2019impact})\normalsize
        \item Public hospitals in Chile improved when older physicians were removed as CEO \scriptsize (\cite{otero2022managers}) \normalsize
    \end{itemize}

    \vspace{5mm}

    \textbf{Hospital Response to Pay-for-Performance}
    \begin{itemize}
        \item Hospitals decreased readmissions by 40\% on average, partially due to selective patient practices \scriptsize (\cite{gupta2021impacts}) \normalsize
        \item Market concentration and systems affect response \scriptsize (\cite{kunz2024assessing}) \normalsize
        \item Reward program had little impact on quality \scriptsize (\cite{us2015hospital}; \cite{norton2018moneyball}; \cite{friedson2019so})
    \end{itemize}
\end{frame}

\begin{frame}{Rest of Talk}

    Setting and Data Collection

    \vspace{4mm}

    About Executive Teams

    \vspace{4mm}

    Theoretical Motivation

    \vspace{4mm}

    Effect of Clinical Leadership

    \vspace{4mm}

    Comparison to For-Profits

    \vspace{4mm}

    Signaling vs. Managing
\end{frame}

\section{Setting and Data Collection}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Setting and Data Collection}
\end{transitionframe}

\begin{frame}{Nonprofit Hospital Executives}\label{nonprofithospexec}
\textbf{What do executives do?}

\vspace{1mm}

\begin{wideitemize}
    \item Board of directors selects executives
    \item Executives handle the day-to-day operations of the hospital \hyperlink{executivejob}{\beamergotobutton{executive job}} 
\end{wideitemize}

\vspace{7mm}

\textbf{Executive teams}

\vspace{1mm}

\begin{wideitemize}
    \item Lots of variation in what executive teams look like
    \item Typically CEO and CFO, sometimes CMO 
    \item I focus on the team as a whole instead of specific executive roles
\end{wideitemize}

\end{frame}

\begin{frame}{}
    \begin{tikzpicture}
        % First image in the top-left
        \node[anchor=north west] at (0, 4) {\includegraphics[width=0.7\textwidth]{Graphics/990_snip_frontpage.PNG}};
        % Second image in the middle
        \node[anchor=north west] (img2) at (2.5, 1.5) {\includegraphics[width=0.75\textwidth]{Graphics/Screenshot 2024-10-03 160258.png}};
        \draw[line width=1mm, eggplant] (img2.north west) rectangle (img2.south east);
    \end{tikzpicture}
\end{frame}

\begin{frame}{Data Collection Process}
    \begin{enumerate}
        \item Get list of Employee Identification Numbers (EIN) that represent nonprofit hospitals

        \vspace{5mm}
        \item Download tax forms from 2009-2015 from Nonprofit Explorer API

        \vspace{5mm}
        \item Use Optimal Character Recognition to extract text from all of these documents

        \vspace{5mm}
        \item Write string cleaning algorithms to extract the correct name, title, position of each entry
    \end{enumerate}
\end{frame}

\begin{frame}{Merge to Other Data}
    \textbf{Hospital:} link EIN and ID number from the American Hospital Association (AHA) Survey based on name and location

    \vspace{1mm}
    \begin{itemize}
        \item \underline{AHA} data for hospital demographics
        \item \underline{Hospital Compare} data for readmission and mortality rates
        \item Center for Medicare and Medicaid Services (CMS) \underline{Case Mix Index file} for patient complexity
        \item \underline{Health Care Cost Report Information System (HCRIS)} for information on pay-for-performance rewards or penalties
    \end{itemize}

    \vspace{7mm}

    \textbf{Physician:} I use executives that self-identify as MD in the tax form data

    \vspace{1mm}

    \begin{itemize}
        \item Manually match physician executives to \underline{NPPES} NPI number
    \end{itemize}
\end{frame}

\begin{frame}{About Pay-for-Performance Policies}
    Policies enacted in 2012 that either penalized for low quality or rewarded for high quality

    \vspace{3mm}

    \textbf{Hospital Readmissions Reduction Program}
    \begin{itemize}
        \item If hospitals have above average readmission rates they get 1-3\% lower reimbursement rates
        \item Pneumonia, Heart Attack (AMI), Heart Failure
    \end{itemize}

    \vspace{3mm}

    \textbf{Hospital Value-Based Purchasing Program}
    \begin{itemize}
        \item safety, efficiency, cost reductions, clinical outcomes, and community engagement, improvement combined into single score metric
        \item ``good" hospitals rewarded
    \end{itemize}

    \begin{block}{\normalsize Takeaway}
        \normalsize Most hospitals had incentive to improve quality to try to avoid meaningful penalties, which were not perfectly predictable
    \end{block}
\end{frame}

\section{About Executive Teams}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Descriptive Statistics}
\end{transitionframe}

\begin{frame}{Executive Teams}
    \begin{table}[ht!]
        \centering
        \begin{tabular}[t]{lccc}
        \toprule
         & Ever & Always & Never\\
         & Clinical Exec & Clinical Exec &  Clinical Exec \\
        \midrule
        Number of Executives & 4.85 & 5.25 & 3.11\\
        \addlinespace[0.2em]
        Fraction Clinical Execs & 0.20 & 0.26 & 0.0\\
        \addlinespace[0.2em]
        Has Clinical CEO & 0.25 & 0.11 & 0.0\\
        \addlinespace[0.2em]
        Has CMO & 0.53 & 0.45 & 0.15\\
        \addlinespace[0.6em]
        Number of Hospitals & 391 & 56 & 445\\
        \bottomrule
        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{Doctors on Executive Teams (N=813)}
    \begin{table}[ht!]
    \centering
    \begin{tabular}[t]{p{6cm} c}
    \toprule
     & Mean\\
    \midrule
    Age & 52.5\\
    Female & 0.11\\
    CEO & 0.16\\
    CMO & 0.31\\
    \addlinespace[0.3em]
    \multicolumn{2}{l}{\textbf{Specialty}}\\
    \hspace{1em}Internal Medicine & 0.32\\
    \hspace{1em}Family Medicine & 0.17\\
    \hspace{1em}Emergency Medicine & 0.11\\
    \hspace{1em}Surgery & 0.09\\
    \hspace{1em}Anesthesiologist & 0.04\\
    \hspace{1em}Other & 0.28\\
    \bottomrule
    \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{Average Team Changes Over Time}
\centering 
    \includegraphics[width=.8\textwidth]{Objects/change_means.pdf}
\end{frame}

\begin{frame}{Changes are not correlated with policy exposure}
\centering \small
\begin{equation*}\label{eq:change2}
    \text{change}_{ht} = \sum_{j=2011}^{2014}\beta_j(\mathbf{1}\{t=j\}\times \text{Program Exposed})_{ht} + \alpha_h + \epsilon_{ht},
    \end{equation*}

    \btVFill

    \includegraphics[width=.7\textwidth]{Objects/change_analysis_plot.pdf}

\end{frame}

\begin{frame}{Hospital Statistics}
\vspace{-2mm}
    \begin{table}[ht!]
    \centering
    \begin{tabular}[t]{lccc}
     & Ever & Always & Never\\
      & Clinical Exec & Clinical Exec & Clinical Exec\\
    \midrule
    \addlinespace[0.3em]
    \multicolumn{4}{l}{\textbf{Hospital Characteristics}}\\
    \hspace{1em}Academic Med. Center & 0.41 & 0.52 & 0.24\\
    \hspace{1em}Number Beds & 231 & 222 & 116\\
    \hspace{1em}Physician Owned & 0.01 & 0.02 & 0.02\\
    \hspace{1em}System Affiliated & 0.58 & 0.52 & 0.44\\
    \addlinespace[0.3em]
    \multicolumn{4}{l}{\textbf{Penalty/Payment Variables}}\\
    \hspace{1em}Ever Received HVBP Incentive & 0.79 & 0.75 & 0.62\\
    \hspace{1em}Penalized for One Condition & 0.18 & 0.20 & 0.13\\
    \hspace{1em}Penalized for AMI + HF/Pneum. & 0.09 & 0.09 & 0.06\\
    \hspace{1em}Penalized for HF + Pneumonia & 0.41 & 0.39 & 0.34\\
    \hspace{1em}Penalized for All Conditions & 0.30 & 0.27 & 0.22\\
    Num. Hospitals & 391 & 56 & 445\\
    \bottomrule
    \end{tabular}
    \end{table}
\end{frame}

\begin{frame}{Raw Outcomes, controlling for baseline case mix index}
\centering
    \includegraphics[width=.9\textwidth]{Objects/outcomes_graph_presentation.pdf}
\end{frame}

\section{Theoretical Framework}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Theoretical Framework}
\end{transitionframe}

\begin{frame}{Hospital Objectives}\label{hosp_obj}
\large
    \begin{equation*}
        \max_{\theta}\hspace{2mm}\alpha\pi(\theta) + (1-\alpha) u(\theta)
    \end{equation*}

    \vspace{5mm} \normalsize

    \begin{wideitemize}
        \item Profit $\pi$ and patient welfare $u$; depend on quality $\theta$
        \item Profit-driven hospitals have high $\alpha$
        \item Abstract away from quality impacting demand for care, price \hyperlink{complex_model}{\beamergotobutton{more complex model}} 
    \end{wideitemize}
\end{frame}

\begin{frame}{Choice of Quality}
Take FOC, rearrange and take derivative wrt $\alpha$:

\vspace{5mm}

\begin{equation*}
    \frac{du'(\theta)}{d\alpha} = \frac{-1}{(1-\alpha)^2}\pi'(\theta)
\end{equation*}

\vspace{5mm}

$\pi'(\theta)$:

\vspace{3mm}

\begin{wideitemize}
    \item In fee-for-service setting: negative $\implies \frac{du'(\theta)}{d\alpha} > 0$
    \item In pay-for-performance setting: positive $\implies \frac{du'(\theta)}{d\alpha} < 0$
\end{wideitemize}

\end{frame}

\begin{frame}{FFS $\implies$ P4P}
    \large
    \begin{align*}
        \frac{d\Delta\theta}{d\alpha}&=\frac{d\theta_{p4p}}{d\alpha}-\frac{d\theta_{ffs}}{d\alpha}\\ 
        &\\
        &=\frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{p4p}}} - \frac{\frac{du'(\theta)}{d\alpha}}{\frac{du'(\theta)}{d\theta_{ffs}}}\\
        &\\
        &> 0
    \end{align*}

    \btVFill \normalsize

    \begin{block}{Takeaway}
        Hospitals with higher weight on profit will respond more to the financial incentives on quality
    \end{block}
\end{frame}

\begin{frame}{Testable Predictions}
    \begin{enumerate}
        \item If non-clinical executives are more profit-driven than clinical executive hospitals, they will reduce quality by more

        \vspace{5mm}
        \item If non-clinical executives are more profit-driven, they will act more like for-profit hospitals

        \vspace{5mm}
        \item Executive teams could reflect an underlying $\alpha$ or they could effectively change $\alpha$
    \end{enumerate}
\end{frame}

\section{Effect of Clinical Leadership on Response}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Effect of Clinical Leadership on Response to P4P}
\end{transitionframe}

\begin{frame}{Estimation}\label{est_equation}
    \textbf{Compare always clinical to never clinical hospital executive teams:}

    \vspace{3mm}\large
    \begin{equation*}
    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation*}

    \vspace{5mm}
    \normalsize
    \begin{wideitemize}
        \item $y_{ht}$: weighted average readmission/ mortality rates
        \item Leverage the \textit{timing} of the programs
    \end{wideitemize}

    \vspace{5mm}

    Different methods:
    \begin{itemize}
        \item No weights
        \item \textbf{``Synthetic DiD" type weights} \hyperlink{synthdid_equation}{\beamergotobutton{weighted}}
        \item Coarsened exact matching \hyperlink{match_equation}{\beamergotobutton{CEM}}
    \end{itemize}
\end{frame}

\begin{frame}{Identification}
\begin{wideitemize}
    \item Trends would have continued in a parallel manner
    \item No other 2012 event correlated with the composition of leadership team and the outcomes
    \item No other hospital characteristics that are perfectly correlated with clinical leadership are driving the results
    \item No anticipation of the incentive changes
\end{wideitemize}
\end{frame}

\begin{frame}{Readmission Rates}\label{main:readrates}
    \begin{columns}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/read_md_nomd_synth_graph.pdf}
    \column{0.5\textwidth}
        \begin{wideitemize}
            \item Non-clinical teams reduce readmissions by .3 ppts more than clinical teams 
            \item Small relative to average readmission rate of 21
            \item Large difference relative to the average effect of these programs (1ppt)
        \end{wideitemize}

        \vspace{7mm}

        \hyperlink{int_margin}{\beamergotobutton{intensive margin}}

    \end{columns}
\end{frame}

\begin{frame}{Mortality Rates}
    \begin{columns}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/mort_md_nomd_synth_graph.pdf}
    \column{0.5\textwidth}
        \begin{wideitemize}
            \item No statistical difference in mortality between clinical and non-clinical teams
        \end{wideitemize}
    \end{columns}
\end{frame}

\section{Comparison to For-Profit Hospitals}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Comparison to For-Profit Hospitals}
\end{transitionframe}

\begin{frame}{Estimation}

    \textbf{Compare both nonprofit leadership team types to the response of for-profits:}

    \vspace{3mm}\large
    \begin{equation*}
    y_{ht} = \beta \text{ (for-profit x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation*}

    \vspace{5mm}
    \normalsize

    Comparison group:
    \begin{itemize}
        \item nonprofit clinical; nonprofit non-clinical
    \end{itemize}

    \vspace{5mm}

    Synthetic diff-in-diff type weights

\end{frame}

\begin{frame}{Readmission Rate}
    \begin{columns}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/fp_read_md_synth_graph.pdf}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/fp_read_nomd_synth_graph.pdf}
    \end{columns}
\end{frame}

\begin{frame}{Mortality Rate}
    \begin{columns}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/fp_mort_md_synth_graph.pdf}
    \column{0.5\textwidth}
        \centering
        \includegraphics[width=.9\textwidth]{Objects/fp_mort_nomd_synth_graph.pdf}
    \end{columns}
\end{frame}

\section{Signaling vs. Managing}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Signaling vs. Managing}
\end{transitionframe}

\begin{frame}{Hypotheses}
    \textbf{Clinical leaders represent underlying $\alpha$}

\vspace{2mm}

    \begin{wideitemize}
        \item Timing of the clinical leader doesn't matter \item Clinical leader represents underlying objectives of the hospital
    \end{wideitemize}

    \vspace{10mm}

    \textbf{Clinical leaders effectively change $\alpha$}

\vspace{2mm}

    \begin{wideitemize}
        \item Timing of the clinical leader does matter
        \item Clinical leaders manage the hospital differently
    \end{wideitemize}
\end{frame}

\begin{frame}{Estimation}
    \begin{figure}[ht!]
\begin{center}

 \begin{tabular}{| m{17em} |}
 \hline
 Signaling:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group:  \hspace{15mm} Never MD \\
 \vspace{2mm}
 Comparison Group: \hspace{3mm} Ever MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{19em}|}
 \hline
 Managing:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{11mm} MD in 2012 \\
 \vspace{2mm}
 Comparison Group:  Ever MD (not in 2012)  \\
 [1ex]
 \hline
 \end{tabular}

\end{center}
 \end{figure}

\end{frame}

\begin{frame}{Results}
    \begin{table}[ht!]
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{\underline{Weighted Avg. Readmission Rate}} & \multicolumn{2}{c}{\underline{Weighted Avg. Mortality Rate}} \\
 & (1) & (2) & (3) & (4)\\
\midrule
Signaling Effect & -0.01 &  & 0.03 & \\
 & (0.07) &  & (0.06) & \\
Managing Effect &  & 0.17 &  & 0.12\\
 &  & (0.1) &  & (0.09)\\
 &  &  &  & \\
Observations & 3,005 & 1,595 & 3,005 & 1,595\\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Conclusion}
    \textbf{We know very little about whether leaders matter in health care}

    \vspace{2mm}

    \begin{wideitemize}
        \item In this paper, I show that non-clinically trained hospital executives internalize financial incentives on quality more than clinically trained executives
        \item Nonprofit objectives are not all created equal, and financial incentives might not be the only way to improve value in health care
    \end{wideitemize}

    \vspace{5mm}

    \textbf{Future Work}

    \vspace{2mm}

    \begin{itemize}
        \item Effects of mandated gender and racial diversity on hospital behavior
        \item Networks of executives over time 
        \item Are leaders correlated with hospital fraudulent behavior?
    \end{itemize}
\end{frame}

\begin{transitionframe}
\centering \Large
    Thank you!

    \vspace{3mm}

    \textcolor{black}{hkagele@emory.edu}
\end{transitionframe}

\section{Appendix}

\begin{transitionframe}
\centering \LARGE
    \textcolor{black}{Appendix Slides}
\end{transitionframe}

\appendix

\begin{frame}{More Complex Model}\label{complex_model}

\btVFill

\hyperlink{hosp_obj}{\beamerreturnbutton{hospital objectives}}

\end{frame}

\begin{frame}{Executive Job}\label{executivejob}

\btVFill

\hyperlink{nonprofithospexec}{\beamerreturnbutton{Nonprofit Hosp Execs}}

\end{frame}

\begin{frame}{Using synthetic DiD weights}\label{synthdid_equation}

\btVFill

\hyperlink{est_equation}{\beamerreturnbutton{main equation}}

\end{frame}

\begin{frame}{Matching}\label{match_equation}

\btVFill

\hyperlink{est_equation}{\beamerreturnbutton{main equation}}

\end{frame}

\begin{frame}{Intensive Margin}\label{int_margin}

    \btVFill

    \hyperlink{main:readrates}{\beamerreturnbutton{main results}}
\end{frame}

\end{document}