### 2025-11-10 Tables/investment_table.tex
add HCRIS investment variables

\begin{table}[ht!]

\caption{\label{tab:investment}Effect of Clinically Trained Executive on Hospital-Level Expenses}
\centering
\begin{tabular}[t]{lccclc}
\toprule
 & Moveable Equip. & Fixed Equipment & Labor Costs & Building Purchases & Total Expenses\\
\midrule
Clinical Exec x Post 2012 & -1 & 6.95 & -1.37 & 0.56 & 1.83\\
 & (2.05) & (6.14) & (2.22) & (7.28) & (8.75)\\
 &  &  &  &  & \\
Observations & 1,532 & 660 & 1,670 & 1,435 & 2,485\\
\bottomrule
\multicolumn{6}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{6}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}