### 2025-11-13 Tables/ceo_table.tex
add table of CEO results

\begin{table}[ht!]

\caption{\label{tab:mainanalysis_ceo}Effect of Clinically Trained CEO on Readmission and Mortality Rates}
\centering
\begin{tabular}[t]{lcc}
\toprule
 & Readmission Rate & Mortality Rate\\
\midrule
Clinical CEO x Post 2012 & -0.75 & -0.4\\
 & (0.45) & (0.43)\\
 &  & \\
Observations & 1,630 & 1,630\\
\bottomrule
\multicolumn{3}{l}{\textsuperscript{} Standard errors are clustered at the}\\
\multicolumn{3}{l}{hospital level.}\\
\multicolumn{3}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*'}\\
\multicolumn{3}{l}{0.05 '.' 0.1}\\
\end{tabular}
\end{table}