### 2025-11-13 Tables/ceo_table.tex
fix footnotes on ceo table

\begin{table}[ht!]

\caption{\label{tab:mainanalysis_ceo}Effect of Clinically Trained CEO on Readmission and Mortality Rates}
\centering
\begin{tabular}[t]{lcc}
\toprule
 & Readmission Rate & Mortality Rate\\
\midrule
Clinical CEO x Post 2012 & -0.75 & -0.4\\
 & (0.45) & (0.43)\\
 &  & \\
Observations & 1,630 & 1,630\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{3}{l}{\rule{0pt}{1em}Standard errors are clustered at the hospital level.}\\
\multicolumn{3}{l}{\rule{0pt}{1em}Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}