### 2025-11-14 Tables/selection_table.tex
add table for patient population instead of graphs

\begin{table}[ht!]

\caption{\label{tab:selection}Effect of Clinically Trained Executive on Patient Population}
\centering
\begin{tabular}[t]{lccc}
\toprule
 & CMI & Frac. Medicare & Frac. Medicaid\\
\midrule
Clinical Exec x Post 2012 & -0.03$^{***}$ & 0 & 0\\
 & (0.01) & (0) & (0.01)\\
 &  &  & \\
Observations & 1,645 & 2,490 & 2,390\\
\bottomrule
\multicolumn{4}{l}{\textsuperscript{} Standard errors are clustered at the hospital level.}\\
\multicolumn{4}{l}{\textsuperscript{} Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1}\\
\end{tabular}
\end{table}